Local Spectral Clustering of Density Upper Level Sets

### 2019-05-20 neurips/local_spectral_density_clustering_monday.tex
added local spectral clustering changes for monday before neurips

%%%%%%%% Local Spectral Clustering of Density Upper Level Sets %%%%%%%%%%%%%%%%%

\documentclass{article}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{float}
\usepackage[export]{adjustbox}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{xr-hyper}
\usepackage{hyperref}
\usepackage[reqno]{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage[parfill]{parskip}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{bm}
\usepackage{fullpage}
\usepackage{mathtools}

\usepackage[square,sort,comma,numbers]{natbib}
\renewcommand{\bibname}{REFERENCES}
\renewcommand{\bibsection}{\subsubsection*{\bibname}}

% Left and right equations
\makeatletter
\newcommand{\leqnomode}{\tagsleft@true\let\veqno\@@leqno}
\newcommand{\reqnomode}{\tagsleft@false\let\veqno\@@eqno}
\makeatother

% Absolute value
\DeclarePairedDelimiter\abs{\lvert}{\rvert}

@@

\title{Local Spectral Clustering of Density Upper Level Sets}
\author{Alden Green, Sivaraman Balakrishnan, and Ryan Tibshirani}
\date{\today}
\maketitle

\begin{abstract}
\vskip 0.1 in % RESOLVED
Spectral clustering methods are a family of popular nonparametric clustering
tools.  Recent works have proposed and analyzed \emph{local} spectral methods,
which extract clusters using locally-biased random walks around a user-specified
seed node.  Several authors have shown that local methods, such as personalized
PageRank (PPR), have worst-case guarantees for certain graph-based measures of
cluster quality.  In contrast to existing works, we analyze PPR in a traditional
statistical learning setup, where we obtain samples from an unknown
distribution, and aim to identify connected regions of high-density (density
clusters).  We introduce two natural criteria for cluster quality, and derive
bounds for these criteria when evaluated on empirical analogues of density
clusters. Moreover, we prove that PPR, run on a neighborhood graph, extracts
sufficiently salient density clusters. Finally, we provide empirical support of our theory.
\end{abstract}

@@

\subsection{Graph Connectivity Criteria}

Here we define a pair of criteria that reflect the quality of a cluster with
respect to $G=(V,E,w)$.  There are many graph-based measures of cluster quality  
that one could consider; see, e.g., \citep{yang2015,fortunato2010} for 
an overview.  The pair of criteria that we focus on are (arguably) quite
natural, and moreover, they play a fundamental role in our analysis 
of the PPR algorithm.  Our two criteria capture the \emph{external} and
\emph{internal} connectivity of a subset $S \subseteq V$, denoted $\Phi(S; G)$
and $\Psi(S; G)$, respectively, and defined below in turn.

\paragraph{External Connectivity: Normalized Cut}

For subsets $S,S' \subseteq V$, we define the cut, degree, and volume functionals as usual,
$$
\cut(S, S'; G) = \sum_{u \in S} \sum_{v \in S'} w_{uv}, \quad \deg(u; G) = \sum_{v \in V} w_{uv}, \quad  \vol(S; G) = \sum_{u \in S} \deg(u;G).
$$
As our notion of external connectivity, we use the \emph{normalized cut}  
of $S$, defined as 
\begin{equation}
\label{eqn: norm_cut}
\Phi(S; G) = \frac{\cut(S; G)}{\min\{\vol(S; G), \vol(S^c; G)\}},
\end{equation} 
where we abbreviate $\cut(S; G) = \cut(S; S^c; G)$.

\paragraph{Internal Connectivity: Inverse Mixing Time}

For $S \subseteq V$, denote by $G[S] = (S, E_S, w_S)$ the subgraph induced by 
$S$ (where the edges are $E_S = E \cap (S \times S)$). Let $\Abf_S,\Dbf_S$
be the adjacency matrix and degree matrix, respectively, of $G[S]$.  Define the
(lazy) random walk matrix as usual, $\Wbf = \frac{\Dbf_S^{-1} \Abf_S + \Ibf{\abs{S} \times \abs{S}}}{2}$, and for $v \in V$,
write 
$$
q_{v}^{(t)}(u) = e_v\Wbf_S^t e_u
$$
for the $t$-step transition probability of a random walk over $G[S]$
originating at $v$.\footnote{Given a starting node $v$ and and a random walk
  defined by transition probability matrix $\mathbf{P}$, the notation $e_v
  \mathbf{P}^t$ is used to denote the distribution of the random walk after $t$
  steps.}   Also write \smash{$\pi = (\pi(u))_{u \in S}$}
for the stationary distribution of this random walk.  (Given the definition of 
$\Wbf_S$, it is well-known that a unique stationary distribution exists and is given by
\smash{$\pi(u) = \deg(u; G[S])/\vol(S; G[S])$}.)

Our internal connectivity parameter will capture the time it takes for the
random walk over $G[S]$ to mix (approach the stationary distribution)
uniformly over $S$.  For this, we first define the \emph{relative pointwise mixing
  time} of $G[S]$ as 
$$
\tau_{\infty}(G[S]) = \min\set{ t: \frac{\pi(u) - q_{v}^{(t)}(u)
	}{\pi(u)} \leq \frac{1}{4}, 
	\; \text{for $u,v \in V$}}. 
$$
%where \smash{$\qbf = (\qbf_v^{(1)}, \qbf_v^{(2)}, ...)_{v \in V}$}, and
%\smash{$\qbf_v^{(m)} = (q_{vu}^{(m)})_{u \in V}$}. 
% RJT: I got rid of the notational dependence on \qbf, because these quantities
% are in turn just defined in terms of the induced subgraph G[S]
% AJG: RESOLVED
Now our internal connectivity parameter is simply the inverse mixing time, 
\begin{equation}
\label{eqn: inv_mixing_time}
\Psi(S; G) = \frac{1}{\tau_{\infty}(G[S])}.
\end{equation}

If $S$ has normalized cut no greater than $\Phi$, and inverse mixing time no
less than $\Psi$, we call it as a \emph{$(\Phi,\Psi)$-cluster}. Both
local \citep{zhu2013} and global \citep{kannan04} spectral algorithms have been
shown to output clusters (or partitions) which approximate the optimal $(\Phi,
\Psi)$-cluster (or partition) for a given graph $G$.\footnote{In the case of
  \citep{kannan04}, the internal connectivity parameter $\phi$ is actually the
  conductance, i.e., the minimum normalized cut within the subgraph $G[S]$. See
  Theorem 3.1 in their paper for details; however, note that $\phi^2 /
  \log(\vol(S)) \leq O(\Psi)$, and so the lower bound on $\phi$ translates to a
  lower bound on $\Psi$.}

@@

For a level $\beta > 0$ and a target volume $\vol_0 > 0$, we define a
\emph{$\beta$-sweep cut} of $\pbf = (p_u)_{u \in V}$ as  
\begin{equation}
\label{eqn: sweep_cuts}
S_\beta = \{u \in V: \frac{p_u}{\Dbf_{uu}} > \frac{\beta}{\vol_{0}}\}.
\end{equation}
Having computed sweep cuts over a range \smash{$\beta \in  
(\frac{1}{40},\frac{1}{11})$},\footnote{The choice of a specific range such as 
\smash{$(\frac{1}{40}, \frac{1}{11})$} is standard in the analysis of PPR
algorithms, see, e.g., \citep{zhu2013}.}
we output a cluster \smash{$\Cest = S_{\beta^*}$}, based on the sweep cut
$S_{\beta^*}$ that minimizes the normalized cut \smash{$\Phi(S_{\beta^*};
  G_{n,r})$} as defined in \eqref{eqn: norm_cut}. For concreteness, we summarize
this procedure in Algorithm \ref{alg: ppr}.

@@

\begin{enumerate}
\item In Section \ref{sec: cluster_quality_bounds}, we derive in Theorem
  \ref{thm: conductance_upper_bound} an upper bound on the normalized cut of a  
  (thickened) empirical density cluster $\Csig[\Xbf]$, under natural geometric 
  conditions (precluding clusters that are arbitrarily thin).

\item Under additional geometric conditions, which exclude sets with large diameter or small bottleneck, we derive in Theorem
  \ref{thm: mixing_time_upper_bound} a lower bound on the
  inverse mixing time of a random walk over $\Csig[\Xbf]$.

\item In Section \ref{sec: consistent_cluster_estimation_with_ppr}, we show in
  Theorems \ref{thm: misclassification_rate} and \ref{thm: consistent_recovery_of_density_clusters} that the bounds on the cluster quality criteria established in Theorems \ref{thm: conductance_upper_bound} and \ref{thm:
    mixing_time_upper_bound} have algorithmic consequences for \ppr. 
  Properly initialized, Algorithm \ref{alg: ppr} has low misclassification error with respect to a small enlargement of the set $\Cset$, and if the density cluster $\Cset$ is particularly well-conditioned, Algorithm \ref{alg: ppr} will perform consistent density cluster estimation in the sense of \eqref{eqn: consistent_density_cluster_recovery}. Corollary \ref{cor: appr} establishes that these statements hold also with respect to an approximate form of \ppr, which can be efficiently computed.

@@

In addition to the background given above, a few related lines of work are worth
highlighting.
Global spectral clustering methods were first developed in the
context of graph partitioning \citep{fiedler1973,donath1973} and their
performance is well-understood in this context (see, e.g.,
\citep{tolliver2006,luxburg2007}).  In a similar vein, several recent works
\citep{mcsherry2001,rohe2011,kamalika2012,balakrishnan2011,lei2015,abbe2018} 
have studied the efficacy of spectral methods in successfully recovering the
community structure in the stochastic block model and variants.

Building on earlier work of
\citep{koltchinskii2000}, \citep{vonluxburg2008,hein2005} studied the limiting behaviour of spectral clustering
algorithms. These authors show that when samples are obtained from a
distribution, and we appropriately construct a geometric graph, the spectrum of
the Laplacian converges to that of the Laplace-Beltrami operator on the
data-manifold. However, relating the partition obtained using the
Laplace-Beltrami operator to the more intuitively defined high-density
clusters can be challenging in general.

@@

Perhaps most similar to our results are the works
\citep{vempala2004,shi2009,schiebinger2015}, who study the consistency of
spectral algorithms in recovering the latent labels in certain parametric and
nonparametric mixture models. These results focus on global rather than local
algorithms, and as such impose global rather than local conditions on the nature
of the density. Moreover, they do not in general ensure recovery of density
clusters, which is the focus in our work.

\section{Cluster Quality Criteria Bounds for Density Clusters}  
\label{sec: cluster_quality_bounds}

\subsection{Geometric Conditions on Density Clusters}

In order to provide meaningful bounds on the normalized cut and inverse mixing
time of an empirical density cluster, we must introduce conditions on the
density $f$. Let $B(x,r) = \{y \in \Rd: \norm{y - x} \leq r\}$ be the closed
ball of radius $r > 0$, centered at $x \in \Rd$.  Given a set $\Aset
\subseteq \Rd$ and $\sigma > 0$, define $\Asig = \Aset + B(0,\sigma) = \{y \in
\Rd: \inf_{x \in \Aset} \norm{y - x} \leq \sigma\}$, which we call the
$\sigma$-expansion of $\Aset$. For a differentiable function $g: \Rd \to \Rd$, write $\nabla g(x)$ to denote the Jacobian of $g$ evaluated at $x \in \Rd$.

We are now ready to give our required conditions, stated with respect to a density
cluster $\Cset \in \Cbb_f(\lambda)$ for some threshold $\lambda > 0$, and an
expansion parameter $\sigma > 0$.

@@

\item 
\label{asmp: low_noise_density}
\emph{Low noise density:} There exists $\gamma,c_0 > 0$ such that for all $x
\in \Rd$ with $0 < \dist(x, \Csig) \leq \sigma$,   
$$
  \inf_{x' \in \Csig} f(x') - f(x) \geq  c_0 \dist(x, \Csig)^{\gamma},
$$
where \smash{$\dist(x,\Aset) = \inf_{x_0 \in \Aset} \norm{x - x_0}$}.

\item
\label{asmp: cluster_separation}
\emph{Cluster separation:}
For all $\Cset' \in \Cbb_f(\lambda)$ with $\Cset' \not= \Cset$,
$$
\dist(\Csig,\Csig') > \sigma,
$$
where \smash{$\dist(\Aset,\Aset') = \inf_{x \in \Aset} \dist(x,\Aset')$}.

\item
\label{asmp: embedding}
\emph{Lipschitz embedding:}
There exists $\mathcal{K} \subseteq \Rd$ convex, and $g: \Reals^d \to \Reals^d$ satisfying, for some $ L \geq 1$,
\begin{equation*}
\det(\nabla g (x) ) = 1,\frac{1}{L}\norm{x - y} \leq \norm{g(x) - g(y)} \leq L \norm{x - y} ~ \text{for all $x,y \in \Reals^d$}
\end{equation*}
such that $\Csig$ is the image of $\mathcal{K}$ by $g$, $\Csig = g(\mathcal{K})$.
Furthermore, there exists $D < \infty$ such that for all $x, x' \in \mathcal{K}$
$$
\norm{x - x'} \leq D.
$$
\end{enumerate}

Note that $\sigma$ plays several roles here, precluding arbitrarily narrow
clusters in \ref{asmp: bounded_density}, flat densities around the level set in \ref{asmp: low_noise_density}, and poorly separated clusters in \ref{asmp: cluster_separation}.

Assumptions \ref{asmp: bounded_density}, \ref{asmp: low_noise_density},
and \ref{asmp: cluster_separation} are used to upper bound $\Phi(\Cset[\Xbf];
G_{n,r})$, whereas \ref{asmp: bounded_density}, and \ref{asmp: embedding} are required to lower bound $\Psi(\Cset[\Xbf]; G_{n,r})$. We note that the
lower bound on minimum density in \ref{asmp: bounded_density} along with \ref{asmp:  
cluster_separation} are similar to the $(\sigma,\epsilon)$-saliency of
\citep{chaudhuri2010}, a standard density clustering assumption, while
\ref{asmp: low_noise_density} is seen in, e.g., \citep{singh2009} (as well as
many other works on density clustering and level set estimation.) While \ref{asmp: embedding} may be less standard, as we will see, it is critical in order to achieve reasonably tight bounds on $\Psi(\Cset[\Xbf];
G_{n,r})$.  It is also worth
highlighting that these assumptions are all local in nature, a benefit of
studying a local algorithm such as \ppr.

%%% AJG 4/29: My attempt to explain why our results aren't fatuous.

We emphasize that while many of these geometric conditions are typical in the density clustering literature, the restrictions we will impose upon them in order to obtain meaningful implications for \pprspace will not be. This is natural. The spectral algorithm we consider is not specifically designed for the task of level set estimation, and in fact one should expect \pprspace to fail to recover -- either in the sense of \eqref{eqn: consistent_density_cluster_recovery}, or indeed any reasonable notion of cluster recovery -- a density cluster of sufficiently large diameter or sufficiently small thickness (though we do not provide any lower bounds to this effect). Indeed, one of the primary motivations of this work was to better understand and characterize the distinctions between those level sets which are well conditioned for spectral algorithms, and those which are not.

In the next several subsections, we will derive bounds on the cluster quality
criteria evaluated on ($\sigma$-expansions of) density clusters. For notational simplicity,
hereafter for $S \subseteq V$, we will abbreviate $\Phi(S; G_{n,r})$ by
$\Phi_{n,r}(S)$, and similarly, $\Psi(S; G_{n,r})$ by $\Psi_{n,r}(S)$, and
$\tau_{\infty}(G_{n,r}[S])$ by $\tau_{n,r}(S)$. We will also use $\nu$ for
Lebesgue measure on $\Rd$, and $\nu_d = \nu(B)$ for the measure of the unit ball
$B=B(0,1)$.

We start with an upper bound on the normalized cut \eqref{eqn: norm_cut} of 
$\Cset_\sigma[\Xbf]$. (In Theorem \ref{thm: conductance_upper_bound}, the upper bound on the density in
Assumption \ref{asmp: bounded_density} will not actually be needed, so we omit
the parameter $\Lambda_\sigma>0$ from the theorem statement.) For $\mathcal{S} \subseteq \Reals^d$ and $r > 0$, let
\begin{equation*}
\pi_{\Pbb,r}(\mathcal{S}) := \frac{\int_{\mathcal{S}} \Pbb(B(x,r)) f(x) dx }{\int_{\Rd} \Pbb(B(x,r)) f(x) dx}.
\end{equation*}

\begin{theorem}
\label{thm: conductance_upper_bound}
Fix $\lambda > 0$, and let $\Cset \in \Cbb_f(\lambda)$ satisfy
Assumptions \ref{asmp: bounded_density}, \ref{asmp: low_noise_density},
and \ref{asmp: cluster_separation}, for some 
$\sigma, \lambda_{\sigma}, c_0, \gamma > 0$. Let $0 < r \leq \sigma/2d$ be such that
\begin{equation}
\label{eqn: weighted_cluster_volume}
\pi_{\Pbb,r}(\Csig) \leq \frac{1}{2}.
\end{equation}
 Then for any $0 < \delta < 1$, $\epsilon > 0$, if
\begin{equation}
\label{eqn: conductance_sample_complexity}
n \geq \frac{(2+\epsilon)^2\log(3/\delta)}{\epsilon^2}\left(\frac{25}
  {6 \lambda_{\sigma}^2\nu(\Csig) \nu_d r^d}\right)^2,
\end{equation}
then
\begin{equation}
  \label{eqn: conductance_additive_error_bound}
  \frac{\Phi_{n,r}(\Csig[\Xbf])}{r} \leq c_1 \frac{d}{\sigma}
  \frac{\lambda}{\lambda_{\sigma}} \frac{(\lambda_{\sigma} -
    c_0\frac{r^{\gamma}}{\gamma+1})}{\lambda_{\sigma}} + \epsilon, 
\end{equation}
with probability at least $1-\delta$ (where $c_1 > 0$ is a universal constant).
\end{theorem}

\begin{remark}
The proof of Theorem \ref{thm: conductance_upper_bound}, along with all other
proofs in this paper, can be found in the supplementary document. The key 
idea is that for any $x \in \Cset$, the simple fact
$B(x,\sigma) \subseteq \Csig$ translates into the upper bound $\nu(\Csig + rB) \leq (1 + 2d r /\sigma)\nu(\Csig)$. We leverage \ref{asmp:
  low_noise_density} to find a corresponding bound on the weighted volume,
then apply standard concentration inequalities to convert from population-
to sample-based results.  
\end{remark}

\begin{remark}
The inequality in \eqref{eqn: conductance_additive_error_bound} is tight in the case of $\Cset = \set{0}$. To see this, let $\Csig = B(0,\sigma)$ and
$$
 f(x)  = \begin{cases}
   \lambda &\text{for $x \in \Csig$}, \\
   \lambda - \dist(x,\Csig)^{\gamma} & \text{ for $0 < \dist(x,\Csig) < r$}, 
\end{cases}
$$
Then, some simple calculations yield
\begin{equation*}
\Ebb(\cut(\Csig[\Xbf]; G_{n,r})) \geq c \lambda \nu_d r^d \Pbb((\Csig + B(0,r)), \quad \textrm{and} \quad \Ebb(\vol_{n,r}(\Csig[\Xbf]; G_{n,r}) \leq c' \lambda \nu_d r^d \Pbb(\Csig)
\end{equation*}
for some constants $c,c' > 0$. Thus the ratio $\Ebb(\cut(\Csig[\Xbf]; G_{n,r})) / \Ebb(\vol_{n,r}(\Csig[\Xbf]; G_{n,r})$ matches \eqref{eqn: conductance_additive_error_bound}, up to constants.

\end{remark}

\subsection{Lower Bound on Inverse Mixing Time}

Next we lower bound the inverse mixing time \eqref{eqn: inv_mixing_time} of $\Csig[\Xbf]$, or equivalently, as $\Psi_{n,r}(\Csig[\Xbf]) = 1/\tau_{n,r}(\Csig[\Xbf])$, we upper bound the mixing time.

\begin{theorem}
	\label{thm: mixing_time_upper_bound}
	Fix $\lambda > 0$, and let $\Cset \in \Cbb_f(\lambda)$ satisfy Assumptions \ref{asmp: bounded_density} and \ref{asmp: embedding} for some $\sigma, \lambda_{\sigma}, \Lambda_{\sigma}, D, K > 0$. Then, for any $0 < r < \sigma/2\sqrt{d}$, with probability one
	\begin{equation}
	\label{eqn: mixing_time_upper_bound}
	\limsup_{n \to \infty}\tau_{n,r}(\Csig[\Xbf]) \leq c_2 \frac{\Lambda_{\sigma}^4 d^3 D^2 L^2}{\lambda_{\sigma}^4 r^2} \log^2\left(\frac{1}{r}\right) + c_3 \log\left(\frac{\Lambda_{\sigma}}{\lambda_{\sigma}}\right)
	\end{equation}
\end{theorem}
for $c_2,c_3 > 0$ universal constants.

Our proof technique involves two key geometric quantities: the \emph{local spread} $s(\widetilde{G}_{n,r})$ (where we abbreviate $\widetilde{G}_{n,r} := G_{n,r}[\Csig[\Xbf]]$ and let $\widetilde{\pi}_{n,r}$ be the stationary distribution over $\widetilde{G}_{n,r}$) and the \emph{conductance} $\widetilde{\Phi}_{n,r}$, defined respectively as
\begin{equation}
\label{eqn: local_spread_and_conductance}
s(\widetilde{G}_{n,r}) := \frac{9}{10} \min_{x \in \Csig[\Xbf]} \set{\deg(x; \widetilde{G}_{n,r}) \cdot \widetilde{\pi}_{n,r}(x) }, \quad \widetilde{\Phi}_{n,r} = \min_{S \subseteq \Csig[\Xbf]} \Phi(S; \widetilde{G}_{n,r}).
\end{equation}

We argue that the random walk over $\widetilde{G}_{n,r}$ quickly escapes sets with stationary distribution less than $s(\widetilde{G}_{n,r})$, and so we avoid a $\log\left(1/\pi_0\right)$ `start penalty' -- where $\pi_0 := \min_{x \in \Csig[\Xbf]} \widetilde{\pi}_{n,r} \lesssim \frac{1}{n}$ -- characteristic to analyses of mixing time, which would render any resultant upper bound on mixing time vacuous. Instead, we obtain the tighter upper bound \footnote{For sequences $a_n, b_n$, we write $a_n \lesssim b_n$ ($a_n \gtrsim b_n$) when there exists $c > 0$ such that $a_n \leq c b_n$ ($a_n \geq c b_n$) for all sufficiently large $n$.}
\begin{equation*}
\tau_{n,r}(\Csig[\Xbf]) \lesssim \frac{1}{\widetilde{\Phi}_{n,r}^2} \log^2\left(1/s(\widetilde{G}_{n,r})\right).
\end{equation*}
Then,
\begin{itemize}
	\item Some straightforward calculations yield $s(\widetilde{G}_{n,r}) \gtrsim \frac{\Lambda_{\sigma} r^d}{\lambda_{\sigma}}$.
	\item To handle the conductance, we introduce a continuous analogue, 
	\begin{equation}
	\label{eqn: continuous_conductance}
	\widetilde{\Phi}_{\Pbb,r} := \min_{\mathcal{S} \subseteq \Csig} \left(\frac{\int_{\Sset} \Pbb(B(x,r) \cap \Sset^c) f(x) dx}{\min \set{\int_{\Sset} \Pbb(B(x,r) \cap \Csig) f(x) dx, \int_{\Sset^c} \Pbb(B(x,r) \cap \Csig) f(x) dx}}\right), \quad \Sset^c = \Csig \setminus \Sset
	\end{equation}
	and show the asymptotic lower bound $\limsup_{n \to \infty} \widetilde{\Phi}_{n,r} \gtrsim \widetilde{\Phi}_{\Pbb,r}$.
	\item Finally, we extend classical isoperimetric results lower bounding $\widetilde{\Phi}_{\Pbb,r}$, when $\Pbb$ is uniform and $\Csig$ convex, to hold under the more general conditions \ref{asmp: bounded_density} and \ref{asmp: embedding}.
\end{itemize}

\begin{remark}
	The embedding assumption \ref{asmp: embedding} and Lipschitz parameter $L$ obviously play an important role in the upper bound of Theorem \ref{thm: mixing_time_upper_bound}. It is clear that there is some interdependence between $L$ and other geometric parameters $\sigma$ and $D$, which might lead one to hope that \ref{asmp: embedding} is non-essential. However, it is not possible to eliminate this condition without incurring an additional factor of at least $(D/\sigma)^d$ in \eqref{eqn: mixing_time_upper_bound}, achieved, for instance, when $\Csig$ consists of two balls of diameter $D$ linked by a cylinder of length $D$ and radius $\sigma$. \citep{abbasi-yadkori2016, abbasi-yadkori2016a} develop theory regarding biLipschitz deformations of convex sets, wherein it is observed that star-shaped sets as well as half-moon shapes of the type we consider in Section \ref{sec: experiments} both satisfy \ref{asmp: embedding} for reasonably small values of $L$.
\end{remark}
\section{Consistent Cluster Estimation}
\label{sec: consistent_cluster_estimation_with_ppr}

@@

%%% AJG 5/8/19: My attempt to explain how to choose these parameters
%%% in practice. Not satisfied with it.

In practice, a reasonable way to choose these hyperparameters is by tuning. For example, if one wanted to successfully recover a density cluster, one could vary each hyperparameter over a grid, retaining outputs $\Cest$ of Algorithm \ref{alg: ppr} only if they recover some $\Cset \in \Cbb_f(\lambda)$ and discarding them otherwise. Then simply return the minimum normalized cut set from those $\Cest$ which were retained. Assuming there existed $\lambda > 0, \Cset \in \Cbb_f(\lambda)$ such that $\Cset$ satisfied the conditions of Theorem \ref{thm: consistent_recovery_of_density_clusters}, and moreover some combination of tuning parameters in the chosen grid satisfied \eqref{eqn: initialization}, this scheme would inherit the consistency guarantees of Theorem \ref{thm: consistent_recovery_of_density_clusters}.

>>>>

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2019

% ready for submission
% \usepackage{neurips_2019}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2019}

% to compile a camera-ready version, add the [final] option, e.g.:
%     \usepackage[final]{neurips_2019}

\usepackage{neurips_2019}

% to avoid loading the natbib package, add option nonatbib:
%     \usepackage[nonatbib]{neurips_2019}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage{microtype}
\usepackage{graphicx}
\usepackage{float}
\usepackage[export]{adjustbox}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{xr-hyper}
\usepackage{hyperref}
\usepackage[reqno]{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage[parfill]{parskip}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{bm}
\usepackage{mathtools}

\DeclarePairedDelimiter\abs{\lvert}{\rvert}

@@

\maketitle

\begin{abstract}
\vskip 0.1 in
  Spectral clustering methods are a family of popular nonparametric clustering
  tools.  Recent works have proposed and analyzed \emph{local} spectral methods,
  which extract clusters using locally-biased random walks around a user-specified
  seed node.  In contrast to existing works, we analyze PPR in a traditional
  statistical learning setup, where we obtain samples from an unknown
  distribution, and aim to identify connected regions of high-density (density
  clusters).  We prove that PPR, run on a neighborhood graph, extracts
  sufficiently salient density clusters, and provide empirical support of our theory.
\end{abstract}

@@

@@

For a level $\beta > 0$ and a target volume $\vol_0 > 0$, we define a
\emph{$\beta$-sweep cut} of $\pbf = (p_u)_{u \in V}$ as  
\begin{equation}
\label{eqn: sweep_cuts}
S_\beta = \{u \in V: \frac{p_u}{\Dbf_{uu}} > \frac{\beta}{\vol_{0}}\}.
\end{equation}

Having computed sweep cuts $S_{\beta}$ over a range \smash{$\beta \in  
(\frac{1}{40},\frac{1}{11})$},\footnote{The choice of a specific range such as 
\smash{$(\frac{1}{40}, \frac{1}{11})$} is standard in the analysis of PPR
algorithms, see, e.g., \citep{zhu2013}.}, we then  output a cluster estimate $\widehat{C} = S_{\beta^*}$ to have minimum normalized cut $\Phi(S_{\beta^{\star}}; G_{n,r})$, where for $S \cup S^c = G_{n,r}$, $\cut(S;G_{n,r}) := |\set{(u,v) \in E : u \in S, v \in S^c}|$, $\vol(S; G_{n,r}) := \sum_{u \in S} \Dbf_{uu}$, and 
\begin{equation}
\label{eqn: normalized_cut}
\Phi(S; G_{n,r}) := \frac{\cut(S;G_{n,r})}{\min \set{\vol(S; G_{n,r}), \vol(S^c; G_{n,r})}}.
\end{equation}
For concreteness, we summarize this procedure in Algorithm \ref{alg: ppr}.

@@

\begin{enumerate}
	\item In Section \ref{sec: consistent_cluster_estimation_with_ppr}, we introduce a set of natural geometric conditions. We formalize a measure of difficulty based on these geometric conditions, and show that when properly initialized, the misclassification error of Algorithm \ref{alg: ppr} is upper bounded by this difficulty measure.

	\item We further show that if the density cluster $\Cset$ is particularly well-conditioned, Algorithm \ref{alg: ppr} will perform consistent density cluster estimation in the sense of \eqref{eqn: consistent_density_cluster_recovery}.

	\item Corollary \ref{cor: appr} establishes that these statements hold also with respect to an approximate form of \ppr, which can be efficiently computed.

	\item In \textcolor{red}{Section 3}, we detail some of the main technical machinery required to prove our main results, highlighting the part various geometric quantities play in the ultimate difficulty of the clustering problem.

@@

In addition to the background given above, a few related lines of work are worth
highlighting. Building on earlier work of
\citep{koltchinskii2000}, \citep{vonluxburg2008,hein2005} studied the limiting behaviour of spectral clustering
algorithms. These authors show that when samples are obtained from a
distribution, and we appropriately construct a geometric graph, the spectrum of
the Laplacian converges to that of the Laplace-Beltrami operator on the
data-manifold. However, relating the partition obtained using the
Laplace-Beltrami operator to the more intuitively defined high-density
clusters can be challenging in general.

@@

More similar to our results are the works
\citep{vempala2004,shi2009,schiebinger2015}, who study the consistency of
spectral algorithms in recovering the latent labels in certain parametric and
nonparametric mixture models. These results focus on global rather than local
algorithms, and as such impose global rather than local conditions on the nature
of the density. Moreover, they do not in general ensure recovery of density
clusters, which is the focus in our work.

\section{Estimation of Well-Conditioned Density Clusters.}
\label{sec: consistent_cluster_estimation_with_ppr}

\subsection{Geometric Conditions on Density Clusters}
\label{subsec: geometric_conditions}

As mentioned previously, successful recovery of a density cluster by \pprspace requires the density cluster to be geometrically well-conditioned. At a minimum, we wish to avoid dumbbell-like sets $\Cset$ which contain an arbitrarily thin bridge, and as in \cite{chaudhuri2010} we therefore introduce a buffer zone around $\Cset$. Letting $B(x,r)$ be the closed ball of radius $r > 0$ centered at $x \in \Rd$, for a given cluster $\Cset \subseteq \Rd$ and $\sigma > 0$, we refer to $\Csig := \set{y \in \Reals^d: \inf_{x \in \Cset} \norm{y - x} \leq \sigma}$ as the $\sigma$-expansion of $\Cset$, and state our conditions with respect to $\Csig$.

More generally, over the neighborhood graph $G_{n,r}$ we would like the empirical cluster $\Csig[\Xbf]$ to be \textcolor{red}{well connected} everywhere in its interior, and \textcolor{red}{poorly connected} to the rest of $\Xbf$.  This intuition motivates our required conditions, stated with respect to a density
cluster $\Cset \in \Cbb_f(\lambda)$ for some threshold $\lambda > 0$, and an
expansion parameter $\sigma > 0$.

@@

	\item
	\label{asmp: cluster_separation}
	\emph{Cluster separation:}
	For all $\Cset' \in \Cbb_f(\lambda)$ with $\Cset' \not= \Cset$,
	$$
	\dist(\Csig,\Csig') > \sigma,
	$$
	where \smash{$\dist(\Cset,\Cset') = \inf_{x \in \Cset} \dist(x,\Cset')$}.

	\item 
	\label{asmp: low_noise_density}
	\emph{Low noise density:} There exists $\gamma,c_0 > 0$ such that for all $x
	\in \Rd$ with $0 < \dist(x, \Csig) \leq \sigma$,   
	$$
	\inf_{x' \in \Csig} f(x') - f(x) \geq  c_0 \dist(x, \Csig)^{\gamma},
	$$
	where \smash{$\dist(x,\Cset) = \inf_{x_0 \in \Cset} \norm{x - x_0}$}.

	%%% AJG 5/20: Should I turn this into two conditions?
	\item
	\label{asmp: embedding}
	\emph{Lipschitz embedding:}
	$\Csig$ is the image of a convex set under a biLipschitz, measure preserving mapping. Formally, there exists $\mathcal{K} \subseteq \Rd$ convex, and $g: \Reals^d \to \Reals^d$ such that $\det(\nabla g (x)) = 1$ for all $x \in \Csig$, and for some $L \geq 1$,
	\begin{equation*}
	\frac{1}{L}\norm{x - y} \leq \norm{g(x) - g(y)} \leq L \norm{x - y} ~ \text{for all $x,y \in \Csig$}
	\end{equation*}
	such that $\Csig$ is the image of $\mathcal{K}$ by $g$, $\Csig = g(\mathcal{K})$.
	Furthermore, there exists $D < \infty$ such that for all $x, x' \in \mathcal{K}$
	$$
	\norm{x - x'} \leq D.
	$$
	\item
	\label{asmp: bounded_volume}
	\emph{Bounded volume:}
	Let the neighborhood graph radius $0 < r \leq \sigma/2d$ be such that
	\begin{equation}
	\label{eqn: weighted_cluster_volume}
	\frac{\int_{\Csig} \Pbb(B(x,r)) f(x) dx }{\int_{\Rd} \Pbb(B(x,r)) f(x) dx} \leq \frac{1}{2}.
	\end{equation}
\end{enumerate}

The cluster separation \ref{asmp: cluster_separation} and low noise density \ref{asmp: low_noise_density} conditions guarantee \textcolor{red}{poor connectivity} between $\Csig[\Xbf]$ and $\Xbf \setminus \Csig[\Xbf]$, whereas \ref{asmp: bounded_density} and \ref{asmp: embedding} ensure high connectivity within $\Csig[\Xbf]$. \textcolor{red}{It may not be immediately obvious how \ref{asmp: embedding} contributes to geometric conditioning. For now, we observe merely that random walks will mix slowly over sets with large diameter, and make some more detailed commentary in Section 3.} Finally, \ref{asmp: bounded_volume} is a relatively harmless technical condition, merely excluding the case where $\Csig$ contains over half the total mass.

@@

In practice it is clearly not feasible to set hyperparameters based on the underlying (unknown) density $f$. Typically, one tunes \pprspace over a range of hyperparameters and optimizes for some criterion such as normalized cut; it is unclear how this scheme would affect the performance of \pprspace in the density clustering context.

\paragraph{Density cluster estimation by \ppr.}
